---
title: '**BANK MARKETING DATA**'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Report on the mining tasks**   

This report is the fruit of endless hours of research in the technics used by the users of R to prepare data, 
indeed it might most certainly not perfect nor as close to accuracy as it should be, however, the advanced coding technics used here are of value, especially the emphasis on the ggplot and the tidyverse data package, as it is a mandatory toolkit for any data scientist.
At the end, looking forward to hear from you any feedback with which i can improve my work.

### **univariate data exploration**  


```{r }
#upload the bank data set
bank <- read.csv2("~/STUDIES/TBS Courses/DM course/raw data/bank.csv")
#view the head of data frame of the set loaded
head(bank)
#structure of the data set
str(bank)
#create a data frame of numerical variables
banknum <- bank[,c("age","balance","day","duration","campaign","pdays","previous")]
#create a data frame of multinomial variables
bankcat <- bank[,c("job","marital","education","contact","month","poutcome")]
#create a data frame of binary variables
bankbin <- bank[,c("default","housing","loan","y")]
```


We will study each set of categories one by one, we will start by the banknum set
We will use the pearson correlation to determine the relation between them, for descriptive statistics we use the function describe from the psych package as it gives us better insights on the normality of the variables and the anaomalies.


```{r}
#Upload the necessary packages
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(corrplot))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(purrr))
```


## Numerical variables exploration
we will also use the function multi.hist to perform a multiple histogram output with density plots. We will also conduct individual normality test, using the shapiro.test function.


```{r }
attach(banknum)
psych::describe(banknum)
rcorr(as.matrix(banknum))
#we chose to not visualize the correlations because there's no significant ones
#prepfereably to visualize with ggplot; rectification required
multi.hist(banknum, bcol="orange",
                    dcol="blue")
shapiro.test(age)
shapiro.test(balance)
shapiro.test(day)
shapiro.test(duration)
shapiro.test(campaign)
shapiro.test(pdays)
shapiro.test(previous)
```


## Categorical variables exploration
now we proceed to analyze the categorical variables in our data set.
we will visualise with barplots and then we will conduct assosiation tests.


```{r }
summary(bankcat)
p1 <-ggplot(data = bankcat, aes(x = job)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
p2 <-ggplot(data = bankcat, aes(x = marital)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3 <-ggplot(data = bankcat, aes(x = education)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
p4 <-ggplot(data = bankcat, aes(x = contact)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
p5 <-ggplot(data = bankcat, aes(x = month)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
p6 <-ggplot(data = bankcat, aes(x = poutcome)) +
     geom_bar(fill="orange", color="black") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1;p2;p3;p4;p5;p6
grid.arrange(p1,p2,p3,p4,p5,p6,
             top="Plot matrix of the categorical variables in the bank data set")
```


## Binary variables explorations
Now with the categorical and numerical variables have been visualized, it's time to proceed to the binary data.
In this section we will have simple barplot of the categories in each binary variable.


```{r }
#summary of the binary data
summary(bankbin)
p11 <-ggplot(data = bankbin, aes(x=default)) + 
            geom_bar(fill="orange", color="black") 
p12 <-ggplot(data = bankbin, aes(x=housing)) + 
            geom_bar(fill="orange", color="black")
p13 <-ggplot(data = bankbin, aes(x=loan)) + 
            geom_bar(fill="orange", color="black")
p14 <-ggplot(data = bankbin, aes(x=y)) + 
            geom_bar(fill="orange", color="black")
grid.arrange(p11,p12,p13,p14,
             top = "Plot matrix of binary data of the bank dataset")
```


## **Bivariate Exploration**
In this section we will try to explore the associations between each variable type, and than the between variable's types
### Categorical 
In this subsection we will construct the contingency tables of possible combinations of the categorical variables, and then we will use these tables to conduct the chi2 test.


```{r }
attach(bankcat)
t1 <-table(job, marital)
t2 <-table(job, education)
t3 <-table(job, contact)
t4 <-table(job, month)
t5 <-table(job, poutcome)
t6 <-table(marital, education)
t7 <-table(marital, contact)
t8 <-table(marital, month)
t9 <-table(marital, poutcome)
t10 <-table(education,contact)
t11 <-table(education,month)
t12 <-table(education,poutcome)
t13 <-table(contact,month)
t14 <-table(contact,poutcome)
t15 <-table(month,poutcome)
tables <- list(t1, t2, t3, t4, t5, t6, t7, t8, t9, t10, t11, t12, t13, t14, t15)
map(tables, chisq.test)
```


### Binary
In this subsection we will explore the inter-binary association within the binary data.
As usual, we construct the contingency tables, and then we conduct the chi2 test.


```{r }
attach(bankbin)
t21 <-table(default, housing)
t22 <-table(default, loan)
t23 <-table(default, y)
t24 <-table(housing, loan)
t25 <-table(housing, y)
t26 <-table(loan, y)
tables1 <- list(t21, t22, t23, t24, t25, t26)
map(tables1, chisq.test)
```


### Binarys vs numerical data
In this subsection we will explore the association between "y", and the rest of the binary variables, this is mainly by conduction boxplots visualization using the ggplot2 package.


```{r }
attach(bank)
a1 <-ggplot(data = bank, aes(x= y, y= age))+ 
     geom_boxplot(fill= "orange", color="black") + 
     coord_flip()
a2 <-ggplot(data = bank, aes(x= y, y= balance))+ 
     geom_boxplot(fill= "orange", color="black") + 
     coord_flip()
a3 <-ggplot(data = bank, aes(x= y, y= day))+ 
     geom_boxplot(fill= "orange", color="black") + 
     coord_flip()
a4 <-ggplot(data = bank, aes(x= y, y= duration))+ 
     geom_boxplot(fill= "orange", color="black")+ 
     coord_flip()
a5 <-ggplot(data = bank, aes(x= y, y= campaign))+ 
     geom_boxplot(fill= "orange", color="black") + 
     coord_flip()
a6 <-ggplot(data = bank, aes(x= y, y= pdays)) + 
     geom_boxplot(fill= "orange", color="black") + 
     coord_flip()
a1 
a2
a3
a4
a5
a6
grid.arrange(a1,a2,a3,a4,a5,a6,
             top="Binary variables difference of groups regarding to 'y'")
```


**"Housing" vs the rest og the numerical variables**


```{r }
attach(bank)
a11 <-ggplot(data = bank, aes(x= housing, y= age)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a21 <-ggplot(data = bank, aes(x= housing, y= balance)) + 
      geom_boxplot(fill= "orange", color= "black") + 
      coord_flip()
a31 <-ggplot(data = bank, aes(x= housing, y= day)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a41 <-ggplot(data = bank, aes(x= housing, y= duration)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a51 <-ggplot(data = bank, aes(x= housing, y= campaign)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a61 <-ggplot(data = bank, aes(x= housing, y= pdays)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a11
a21
a31
a41
a51
a61
grid.arrange(a11,a21,a31,a41,a51,a61,
             top= "Numerical data and their propreties comparing to housing")
```


**"Loan" vs the rest of the numerical variables**


```{r }
attach(bank)
a12 <-ggplot(data = bank, aes(x= loan, y= age)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a22 <-ggplot(data = bank, aes(x= loan, y= balance)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a32 <-ggplot(data = bank, aes(x= loan, y= day)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a42 <-ggplot(data = bank, aes(x= loan, y= duration)) + 
  geom_boxplot(fill= "orange", color="black") + 
  coord_flip()
a52 <-ggplot(data = bank, aes(x= loan, y= campaign)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a62 <-ggplot(data = bank, aes(x= loan, y= pdays)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a12
a22
a32
a42
a52
a62
grid.arrange(a12,a22,a32,a42,a52,a62,
             top="numerical variables and their propreties comparing to 'loan'")
```


**"default" vs the rest of the categorical variables**


```{r }
attach(bank)
a13 <-ggplot(data = bank, aes(x= default, y= age)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a23 <-ggplot(data = bank, aes(x= default, y= balance)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a33 <-ggplot(data = bank, aes(x= default, y= day)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a43 <-ggplot(data = bank, aes(x= default, y= duration)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a53 <-ggplot(data = bank, aes(x= default, y= campaign)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a63 <-ggplot(data = bank, aes(x= default, y= pdays)) + 
      geom_boxplot(fill= "orange", color="black") + 
      coord_flip()
a13
a23
a33
a43
a53
a63
grid.arrange(a13,a23,a33,a43,a53,a63,
             top ="the variable default in comparision to numerical variables")
```


### Numerical variables and the responses in "y"
In this subsection we will explore the numerical variables pair-wise while discriminating them according to the response in the variable "y"


**"age" vs the rest of the numerical variables**


```{r }
attach(bank)
b11 <-ggplot(data = bank, aes( y= age))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b21 <-ggplot(data = bank, aes( y= age))+
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b31 <-ggplot(data = bank, aes( y= age))+ 
      geom_boxplot(aes(fill = y)) +
      facet_grid(education~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b41 <-ggplot(data = bank, aes( y= age))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b51 <-ggplot(data = bank, aes( y= age))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b61 <-ggplot(data = bank, aes( y= age))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b11
b21
b31
b41
b51
b61
```


**"balance" vs the rest of the numerical variables** 
 
 
```{r }
b12 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b22 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b32 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(education~.) +
      coord_flip() +
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b42 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) + 
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b52 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b62 <-ggplot(data = bank, aes( y= balance)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b12
b22
b32
b42
b52
b62
```


**"day" vs the rest of the numerical variables**
 
 
```{r }
b13 <-ggplot(data = bank, aes( y= day)) +
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b23 <-ggplot(data = bank, aes( y=day)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b33 <-ggplot(data = bank, aes( y= day))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(education~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b43 <-ggplot(data = bank, aes( y= day)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b53 <-ggplot(data = bank, aes( y= day)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b63 <-ggplot(data = bank, aes( y= day)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1)) 
b13
b23
b33
b43
b53
b63
```


**"duration" vs the rest of the numerical variables**


```{r }
b14 <-ggplot(data = bank, aes( y= duration)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b24 <-ggplot(data = bank, aes( y= duration))+ 
      geom_boxplot(aes(fill = y)) + facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b34 <-ggplot(data = bank, aes( y= duration)) +  
      geom_boxplot(aes(fill = y)) +  
      facet_grid(education~.) +
      coord_flip()+
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b44 <-ggplot(data = bank, aes( y= duration)) + 
      geom_boxplot(aes(fill = y)) +  
      facet_grid(contact~.) + 
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b54 <-ggplot(data = bank, aes( y= duration)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b64 <-ggplot(data = bank, aes( y= duration)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() +
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b14
b24
b34
b44
b54
b64
```


**"campaign" vs the rest of the numerical variables**


```{r }
b15 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b25 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b35 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(education~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b45 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) + 
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b55 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b65 <-ggplot(data = bank, aes( y= campaign)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) + 
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b15
b25
b35
b45
b55
b65
```
 
 
**"pdays" vs the rest of the numerical variables**


```{r }
b16 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b26 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b36 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(education~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b46 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b56 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) +
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b66 <-ggplot(data = bank, aes( y= pdays)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b16
b26
b36
b46
b56
b66
```


**"previous" vs the rest of the numerical variables**


```{r}
b17 <-ggplot(data = bank, aes( y= previous)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(job~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b27 <-ggplot(data = bank, aes( y= previous))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(marital~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b37 <-ggplot(data = bank, aes( y= previous))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(education~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b47 <-ggplot(data = bank, aes( y= previous))+ 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(contact~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b57 <-ggplot(data = bank, aes( y= previous)) + 
      geom_boxplot(aes(fill = y)) + 
      facet_grid(month~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b67 <-ggplot(data = bank, aes( y= previous)) +
      geom_boxplot(aes(fill = y)) + 
      facet_grid(poutcome~.) +
      coord_flip() + 
      theme(strip.text.y = element_text(angle = 360, hjust = 1))
b17
b27
b37
b47
b57
b67
```

